---
title: "VeloxRaster Objects"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

<br>

## The `VeloxRaster` Class

Most `velox` functionality is implemented as methods of the `VeloxRaster` class.
Objects of class `VeloxRaster` are [R5 (Reference Class)](http://adv-r.had.co.nz/R5.html) objects and are thus mutable. 
Hence, the usual copy on modify semantics do not apply. 
Rather, calling a raster-modifying `VeloxRaster` method, e.g. the crop method (`vx$crop()`), will modify the object the method is applied to directly. 
The main advantage of having mutable raster objects is that it saves memory: Unless expressely requested by the user, the raster data will not be copied for modification.

As an example, consider the following code:

```{r message=FALSE} 
require(velox)

## Create VeloxRaster from matrix
mat <- matrix(1:100, 10, 10)
vx <- velox(x = mat, extent = c(0,10,0,10), res = c(1,1))

## Crop VeloxRaster
vx$crop(c(0,5,0,5))
print(vx$dim)
```

Due to mutability, the `vx` object is permanently modified by the crop operation.
If we wish to keep a copy of the original `vx` object, we have to instruct R to make a copy explicity:

```{r message=FALSE} 
## Create VeloxRaster from matrix
mat <- matrix(1:100, 10, 10)
vx <- velox(x = mat, extent = c(0,10,0,10), res = c(1,1))

## Copy the VeloxRaster
vx.copy <- vx$copy()

## Crop the copied VeloxRaster
vx.copy$crop(c(0,5,0,5))
print(vx.copy$dim)
print(vx$dim)
```


<br>

## Creating `VeloxRaster` objects

`VeloxRaster` objects are created with the `velox` function.

velox is fully interoperable with the `raster` package; creating `VeloxRaster` objects from `Raster*` is simple:
```{r message=FALSE} 
require(raster)

## Create VeloxRaster from RasterLayer
mat <- matrix(1:100, 10, 10)
ras <- raster(x = mat)
ras.vx <- velox(ras)

## Create VeloxRaster from RasterStack
ras1 <- ras2 <- raster(x = mat)
stk <- stack(ras1, ras2)
stk.vx <- velox(stk)

## Create VeloxRaster from RasterBrick
brk <- brick(ras1, ras2)
brk.vx <- velox(brk)
```

We can also create `VeloxRaster` objects from matrices, or lists of matrices:
```{r message=FALSE} 
## Create VeloxRaster from a matrix
mat <- matrix(1:100, 10, 10)
mat.vx <- velox(mat, extent = c(0,10,0,10), res = c(1,1))

## Create VeloxRaster from a list of matrices
mat.ls <- list(mat, mat)
matls.vx <- velox(mat.ls, extent = c(0,10,0,10), res = c(1,1))
```

Finally, we can read [GDAL-supported](http://www.gdal.org/formats_list.html) raster files from disk:
```{r message=FALSE} 
## Create a RasterLayer
mat <- matrix(1:100, 10, 10)
ras <- raster(x = mat)

## Write it to disk as a GeoTiff file
writeRaster(ras, file.path(tempdir(), 'test.tif'))

## Read the GeoTiff as VeloxRaster
tif.vx <- velox(file.path(tempdir(), 'test.tif'))
print(tif.vx$dim)

## Clean up
unlink(file.path(tempdir(), 'test.tif'))
```
<br>

## Casting `VeloxRaster` objects

Casting `VeloxRaster` objects as `Raster*` objects is also simple:

```{r message=FALSE} 
## Create VeloxRaster from a list of matrices
mat.ls <- list(mat, mat)
vx <- velox(mat.ls, extent = c(0,10,0,10), res = c(1,1))

## Cast first band as RasterLayer
ras <- vx$as.RasterLayer(band = 1)

## Cast as RasterStack
stk <- vx$as.RasterStack()

## Cast as RasterBrick
brk <- vx$as.RasterBrick()
```

We can also cast raster bands as matrices:
```{r message=FALSE} 
## Cast first raster band as matrix
mat <- vx$as.matrix(band = 1)
```
<br>

## Saving `VeloxRaster` objects

Finally, we can save `VeloxRaster` objects to disk as GeoTiff files:
```{r message=FALSE} 
## Create VeloxRaster from a list of matrices
mat.ls <- list(mat, mat)
vx <- velox(mat.ls, extent = c(0,10,0,10), res = c(1,1))

## Save to disk
vx$write(path = file.path(tempdir(), 'test.tif'))

## Clean up
unlink(file.path(tempdir(), 'test.tif'))
```




