---
title: "Extract"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

velox offers fast methods for extracting and summarizing raster data at the locations of spatial objects, that is, polygons, lines, or points.


<br>

## Get raster data

In the following, we demonstrate velox's extract functionality using some of the nightlight emission data available in the [PRIO GRID](http://grid.prio.org) data set. 
The [pgrid R package](https://github.com/hunzikp/pgrid) provides a convenient way to return PRIO GRID data in `VeloxRaster` format:

```{r ch1} 
library(pgrid)
prio.ls <- getPrioRaster(names = 'nlights_calib_mean', years = c(2010:2012))
prio.vx <- prio.ls$raster
priometa.df <- prio.ls$meta
```

Let's plot the 2010 nightlight data:

```{r ch2, message = FALSE} 
library(raster)
nl10.ras <- prio.vx$as.RasterLayer(priometa.df$band[priometa.df$year == 2010])
plot(nl10.ras)
```

<br>

## Polygon extract

We first illustrate value extraction using polygon data. To this end, we use the territory of souvereign states in 2010, as provided through the `cshapes` package:

```{r ch3, message = FALSE} 
library(cshapes)
world.spdf <- cshp(as.Date('2010-01-01'))
```

Now let's extract all raster values for a single country, e.g. Syria, using the `VeloxRaster_extract` method.

```{r ch4, message = FALSE} 
syria.spdf <- world.spdf[world.spdf$CNTRY_NAME == 'Syria',]
extr.ls <- prio.vx$extract(sp = syria.spdf)
head(extr.ls[[1]])
```
By default, `VeloxRaster_extract` returns a list with one element per entry in the input `Spatial*` data frame, and each element consisting of a matrix of raster values. Here the matrix has three columns because `prio.vx` consists of three raster bands.

<br>

## Aggregation functions

Often we don't want 'raw' raster values, but summaries thereof. 
`VeloxRaster_extract` permits passing an R function for summarizing the raster values before returning them.
In the following, we request the arithmetic mean. Note that we set `na.rm = TRUE` in order to handle `NA` values in the raster data.

```{r ch5, message = FALSE} 
nlmean.mat <- prio.vx$extract(sp = syria.spdf, fun = function(x) mean(x, na.rm = TRUE))
print(nlmean.mat)
```

<br>

## The small option

By default, `VeloxRaster_extract` determines whether a raster cell intersects with a polygon by testing whether its center intersects with the polygon. 
For small or oddly shaped polygons, this may lead to `VeloxRaster_extract` returning `NA` values. 
Consider the case of Monaco for illustration:

```{r ch6, message = FALSE, warning = FALSE}
library(rgeos)
monaco.spdf <- world.spdf[world.spdf$CNTRY_NAME == 'Monaco',]
plot(gEnvelope(gBuffer(monaco.spdf, width = 0.25)))  # Set plotting area
plot(prio.vx$as.RasterLayer(1), add = TRUE)
plot(monaco.spdf, add = TRUE)
```

Here, running `VeloxRaster_extract` returns `NA`:

```{r ch7, message = FALSE, warning = FALSE}
monaco.mat <- prio.vx$extract(sp = monaco.spdf, fun = function(x) mean(x, na.rm = TRUE))
print(monaco.mat)
```

To avoid this problem we set `small = TRUE` in `VeloxRaster_extract`. 
Doing so forces velox to look for full polygon-raster-cell overlaps for those polygons that do not intersect with a raster centroid:

```{r ch8, message = FALSE, warning = FALSE}
monaco.mat <- prio.vx$extract(sp = monaco.spdf, fun = function(x) mean(x, na.rm = TRUE), small = TRUE)
print(monaco.mat)
```

<br>

## Line extraction

Raster value extraction for lines works equivalently to the polygon case.
Note, however, that for line geometries, velox always tests for intersections with the full raster-cells, not just the centroids.
As an illustration, consider the following SpatialLinesDataFrame, which consists one line per parallel (i.e. one line at latitude 89.5, one at 88.5, etc). 

```{r ch9, message = FALSE, warning = FALSE}
## Make latitude lines
latVec <- seq(from = 89.5, to = -89.5, by = -1)
Lns.ls <- list()
for (lat in latVec)  {
  coords <- matrix(c(-180, lat, 
                     180, lat), 2, 2, byrow = TRUE)
  Lns <- Lines(list(Line(coords)), paste(lat))
  Lns.ls[[length(Lns.ls)+1]] <- Lns
}
lat.sldf <- SpatialLinesDataFrame(SpatialLines(Lns.ls), data.frame(id = 1:length(Lns.ls)), match.ID = F)
```

We now extract raster values from a 1x1 decimal degree version of PRIO GRID using these lines.

```{r ch10, message = FALSE, warning = FALSE}
## Aggregate PRIO GRID
prio.agg.vx <- prio.vx$copy()
prio.agg.vx$aggregate(factor = 2, aggtype = 'sum')

## Extract raster values
lineval.ls <- prio.agg.vx$extract(sp = lat.sldf)
print(dim(lineval.ls[[1]]))
```

The returned matrices have 360 rows because the intersect with all 180*2 raster cells on each parallel.
Incidentally, we can use these data to create a neat line map (as inspired by [this blog post](http://spatial.ly/2014/08/population-lines/)).

```{r ch11, message = FALSE, warning = FALSE, fig.height = 7, fig.width = 14}

## Raster cell x vector
xVec <- seq(from = -179.5, to = 179.5, by = 1)

## Get a list of only 2010 nightlight data
nl2010.ls <- lapply(lineval.ls, function(x) x[,1])

## Get the minimal nightlight value observed (so we can normalize)
nl.min <- min(unlist(lapply(nl2010.ls, function(x) min(x, na.rm = TRUE))))

## Empty plot
par(mar = c(0,0,0,0), bg = grey(0.25), xaxs = "i", yaxs = "i")
plot(1, type="n", xlab="", ylab="", xlim=c(-180,180), ylim=c(-90, 90), axes=FALSE, bty="n")

## Iterate over parallels
for (i in 1:length(nl2010.ls)) {
  
  ## Get nightlights at this parallel
  val <- nl2010.ls[[i]] - nl.min
  
  ## Replace NAs with zero, multiply with constant (to make variance more visible)
  val[is.na(val)] <- 0
  val <- val * 5
  
  # Mark (non-zero) segments
  counter <- 1
  segmentVec <- rep(0, length(val))
  segmentVec[1] <- ifelse(val[1] == 0, 0, counter)
  for (j in 2:length(val)) {
    if (val[j-1] == 0 & val[j] != 0) {
      counter <- counter+1
      segmentVec[j] <- counter
    } else if (val[j] != 0) {
      segmentVec[j] <- counter
    }
  }
  
  # Add lat to val
  val <- val + latVec[i]
  
  # Plot full parallel in grey
  points(xVec, val, type = 'l', lwd = 1, col = grey(0.5))
  
  # Plot polygon for each segment
  segments <- unique(segmentVec[segmentVec != 0])
  if (length(segments) > 0) {
    for (s in segments) {
      thisXVec <- xVec[segmentVec == s]
      thisYVec <- val[segmentVec == s]
        
      # Plot polygon
      if (length(thisXVec) > 1) {
        polygon(c(min(thisXVec)-1, thisXVec, max(thisXVec)+1),
                c(latVec[i], thisYVec, latVec[i]),
                col=grey(0.25), border = NA)
      }
      
      # Plot polygon contour in white
      points(c(min(thisXVec)-1, thisXVec, max(thisXVec)+1), 
             c(latVec[i], thisYVec, latVec[i]), 
             type = 'l', lwd = 1, col = 'white')
    }
  }
}

```

<br>

## Point extraction

Extracting raster values at the location of points is possible using the `VeloxRaster_extract_points` method.
As a demonstration, let's create some random point data:

```{r ch12, message = FALSE, warning = FALSE}
## Some random point data
set.seed(0)
points.sp <- SpatialPoints(cbind(runif(100, -180, 180), runif(100, -90, 90)))
```

Extracting the respective raster values:
```{r ch13, message = FALSE, warning = FALSE}
## Extract raster values at points
pval.mat <- prio.vx$extract_points(sp = points.sp)
print(dim(pval.mat))
```

<br>

## Using sf objects

The `VeloxRaster_extract` method is fully interoperable with the [sf package](https://cran.r-project.org/package=raster):

```{r ch14, message = FALSE} 
require(sf)

## Cast syria.spdf as sf object
syria.sf <- st_as_sf(x = syria.spdf)

## Use extract as before
extr.ls <- prio.vx$extract(sp = syria.sf)
head(extr.ls[[1]])
```

Similarly, the `VeloxRaster_extract_points` method also works on `sf` objects storing `POINT` data:

```{r ch15, message = FALSE} 
## Cast random points as sf object
points.sf <- st_as_sf(x = points.sp)

## Use extract_points as before
pval.mat <- prio.vx$extract_points(sp = points.sf)
print(dim(pval.mat))
```

`VeloxRaster_extract_points` does not (yet) work with sf data of type `MULTIPOINT`, however.

