---
title: ""
---

<br>


![](logo.svg)

<br>

# Fast raster extraction and manipulation in R

velox is an R package for performing fast extraction and manipulation operations on geospatial raster data. 
velox is fast because:

* All raster operations are performed in C++.
* Geometric operations are implemented with the [Boost Geometry](http://www.boost.org/doc/libs/1_65_1/libs/geometry/doc/html/index.html) libraries.
* All data is held in memory.

velox is fully interoperable with the [raster](https://cran.r-project.org/package=raster), [sp](https://cran.r-project.org/package=raster), and [sf](https://cran.r-project.org/package=raster) packages.


[![Travis-CI Build Status](https://travis-ci.org/hunzikp/velox.svg?branch=master)](https://travis-ci.org/hunzikp/velox)
[![codecov](https://codecov.io/gh/hunzikp/velox/branch/master/graph/badge.svg)](https://codecov.io/gh/hunzikp/velox)
[![CRAN Version](http://www.r-pkg.org/badges/version/velox)](https://cran.r-project.org/package=velox)
`r require(badger); badge_devel("hunzikp/velox", "green")`
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/last-week/velox)](https://www.r-pkg.org/pkg/velox)

<br>

## Optimized for speed

``` {r, echo = FALSE, warning = FALSE, message = FALSE}
## Dependencies
library(velox)
library(sf)
library(raster)
library(rasterVis)

## Load data
nl.vx <- velox('nl.tif')
sui.sf <- readRDS(file = 'sui.Rda')

## Extract using vx
vx.time <- system.time(ex.mat <- nl.vx$extract(sp = sui.sf, fun = function(x) sum(x, na.rm = TRUE)))

## Extract using raster
sui.sp <- as(st_zm(sui.sf), 'Spatial')
#ras.time <- system.time(extract(x = nl.vx$as.RasterLayer(), sui.sp))
ras.time <- c(0,0,54.759)
```
``` {r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 2.75, fig.width = 4.5}
## Plot nightlights
lattice.options(axis.padding=list(factor=0.5))
nightPalette <- colorRampPalette(c("black", 
                                   rgb(4,6,84, maxColorValue=255),
                                   rgb(218,165,32, maxColorValue=255), 
                                   "white"), bias = 0.95)
par(mar = c(0,0,0,0))
image(log(nl.vx$as.RasterLayer()), col = nightPalette(25), axes = FALSE)
```
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 2.75, fig.width = 4.5}
## Plot geometries
ar <- st_area(st_zm(sui.sf))
lnl <- log(as.vector(ex.mat)/as.vector(ar))  # Log nl per area
nightPalette <- colorRampPalette(c(rgb(4,6,84, maxColorValue=255),
                                   rgb(218,165,32, maxColorValue=255), 
                                   "white"), bias = 1)
col <- nightPalette(25)[as.numeric(cut(lnl, breaks = 25))]
par(bg = grey(0.25), mar = rep(0, 4), xaxs = "i", yaxs = "i")
plot(sui.sp, col = col, border = grey(0.1), lwd = 0.1)
```


Aggregating September 2017 [nightlight emissions](https://www.ngdc.noaa.gov/eog/viirs/download_dnb_composites.html) for 2268 Swiss communes takes **`r round(vx.time[3], 2)`** seconds in velox and **`r round(ras.time[3], 2)`** seconds in [raster](https://cran.r-project.org/package=raster).

<br>

## Features

velox currently offers the following features:

* Raster value extraction given *polygons*, *lines*, or *points*
* Focal value calculation (i.e. moving window filters)
* Rasterization of polygons or lines
* Raster aggregation
* Cropping
* Image patch flattening and reconstruction


