---
title: "Installation"
output:
  html_document:
    toc: false
    toc_depth: 2
    toc_float:
      collapsed: false
---

<br>

## System dependencies

velox only requires a C++11 capable compiler, which shouldn't be an issue on most contemporary systems.
However, velox depends on the [rgdal](https://cran.r-project.org/package=rgdal), [rgeos](https://cran.r-project.org/package=rgeos), and [sf](https://cran.r-project.org/package=sf) packages, which require the external GDAL (>= 2.0.0), GEOS (>= 3.3.0), PROJ.4 (>= 4.8.0), and UDUNITS libraries.
On Ubuntu (>= 14.04), install these dependencies using:
```shell
sudo add-apt-repository -y ppa:ubuntugis/ubuntugis-unstable
sudo apt-get update
sudo apt-get install libgdal-dev libproj-dev libgeos-dev libudunits2-dev
```

See the [sf README](https://r-spatial.github.io/sf/) for information about installation on other platforms.

<br>

## R installation

To install the CRAN version ([![CRAN Version](http://www.r-pkg.org/badges/version/velox)](https://cran.r-project.org/package=velox)), use:
```{r eval=FALSE} 
install.packages('velox')
```

To install the development version (`r require(badger); badge_devel("hunzikp/velox", "green")`) from Github:
```{r eval=FALSE} 
require(devtools)
devtools::install_github('hunzikp/velox')
```




