---
title: "Quick Start"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

No time to read the docs? Here's the basics.
See the Usage section in the menu for more detailed documentation of velox functionality.

<br>

## Cast as/from `VeloxRaster`

Create a `VeloxRaster` object from a `raster*` object:

```{r message=FALSE} 
require(raster)
require(velox)

## Create RasterStack
set.seed(0)
ras.ls <- lapply(1:2, function(x) raster(matrix(rnorm(100), 10, 10), xmn=0, xmx=10, ymn=0, ymx=10))
stk <- stack(ras.ls)

## Cast as VeloxRaster
vx <- velox(stk)
```

Or load a `VeloxRaster` object directly from disk:

```{r} 

## Temporary path
temp.path <- file.path(tempdir(), 'test.tif')

## Save as GeoTiff
vx$write(path = temp.path)

## Read from disk
vx <- velox(temp.path)

## Clean up
unlink(temp.path)
```

Cast a `VeloxRaster` object as a `RasterStack`:
```{r} 
## As RasterStack
stk <- vx$as.RasterStack()
```

<br>

## Manipulate `VeloxRaster`

Crop a `VeloxRaster`; note that `VeloxRaster` objects are *mutable*:

```{r} 

## Copy the VeloxRaster and check its extent
vx_copy <- vx$copy()
as.vector(vx_copy$extent)

## Crop and check extent again
vx_copy$crop(c(0,5,0,5)) # Crop by extent; format c(xmin, xmax, ymin, ymax)
as.vector(vx_copy$extent)
```

Aggregate a `VeloxRaster`:

```{r} 

## Copy the VeloxRaster and aggregate by factor 2
vx_copy <- vx$copy()
vx_copy$aggregate(factor = 2, aggtype = 'sum')
```


Apply a moving window filter:


```{r} 

## Copy the VeloxRaster and apply a moving window filter to the first raster band
vx_copy <- vx$copy()
vx_copy$meanFocal(weights = matrix(1, 3, 3), bands = 1)
```

<br>

## Extract raster values

Extract raster values given a set of polygons:

```{r message = FALSE} 
## Make SpatialPolygonsDataFrame
library(sp)
library(rgeos)
set.seed(0)
extent <- as.vector(vx$extent)
coords <- cbind(runif(5, extent[1], extent[2]), runif(10, extent[3], extent[4]))
sp <- SpatialPoints(coords)
spol <- gBuffer(sp, width=2, byid=TRUE)
spdf <- SpatialPolygonsDataFrame(spol, data.frame(id=1:length(spol)), FALSE)

## Extract values and calculate mean
ex.mat <- vx$extract(spdf, fun=mean)
print(ex.mat)

## Also works with sf objects
library(sf)
poly.sf <- st_as_sf(spdf)
ex.mat2 <- vx$extract(poly.sf, fun=mean)
all(ex.mat == ex.mat2)
```

<br>

## Rasterize polygons

Rasterize some polygons:

```{r message = FALSE} 
## Copy VeloxRaster
vx_copy <- vx$copy() 

## Rasterize polygons using "id" column
vx_copy$rasterize(spdf, field="id", band=1, background = 0)
```

<br>

## Image patch flattening

Transform the content of a raster into a matrix of 'flat' patches:

```{r message = FALSE} 
## Get flattened raster image
flat.mat <- vx$im2col(wrow = 3, wcol = 3, band = 1, rowframe = 0, colframe = 0)
print(dim(flat.mat))

## Replace 'center' pixels with zeros
flat.mat[,5] <- 0

## Copy flattened image to 2nd band of raster
vx$col2im(mat = flat.mat, wrow = 3, wcol = 3, band = 2, rowframe = 0, colframe = 0)

```

<br>

## Help

Get help for e.g. the extract method:

```{r message = FALSE} 
?VeloxRaster_extract
```